### 2025-02-02 products/manuscript/Manuscript.qmd
fixed some typo

---
title: "The effects of multi environmental exposure on Tuberculosis Incidence in Asian and African countris"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

The structure below is one possible setup for a manuscript, or a general
data analysis project (including the course project). Adjust as needed.
You don't need to have exactly these sections, but the content covering
those sections should be addressed.

This uses MS Word as output format. [See
here](https://quarto.org/docs/output-formats/ms-word.html) for more
information. You can switch to other formats, like html or pdf. See [the
Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

-   Muhammad Nasir$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
-   Second Author$^{3, *}$
-   Third Author$^{3, \dagger}$
-   Last Author$^{1, 2, \land}$

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.
2.  Faculty of Public Health, Universitas Tadulako, Palu, Indonesia.
3.  Yet another Affiliation.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: some\@email.com

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

Tuberculosis is old infectious disease which still exists and become
major public health in the world, particularly in developing countries.
Tuberculosis is linked to environmental factors. This study will analyze
the effects of multi environmental factors on tuberculosis incidence in
Asia and Africa regions, where the majority of tuberculosis coming from
these regions. The environmental factors and country economic and human
development index as covariates.

## Description of data and data source

The study will use multi sources of data including World Bank, World
Health Organization (WHO), United Nation for Children Fund (Unicef),
UN-HABITAT, Emissions Database for Global Atmospheric Research (EDGAR),
and Institute for Health Metrics and Evaluation (IHME). Different
dataset contains different variable, however all dataset have same year,
country name and country codes that can be used to combine the dataset.
The data also contain information of countries globally from 2000 to
2020. The variables will include proportion of urban population living
in slum areas, percentage of population using safely managed sanitation
services, percentage of population using at least basic drinking water
service, fuel for cooking, Population-weighted exposure to ambient PM2.5
pollution, total green houses, and will include GDP/ country economic
status, human development index, and Universal Health Coverage as
covariates.

Note: I have downloaded the dataset, but I have not conducted any
cleaning or exploration on the dataset. List of data sources and website
is provided in an excel fine in my data folder. I will include them in
my reference for the next steps.

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

The main aim of this study is to look at the impacts of environmental
factors on tuberculosis incidence.

RQ: What environmental factors associated with tuberculosis incidence?

H0: there is no relationship between proportion of urban population
living in slum areas and tuberculosis.

H0: thee is no relationship between safely managed sanitation and
tuberculosis incidence

H0: there is no relationship between drinking water and tuberculosis
incidence

H0: there is no relationship between fuel for cooking and tuberculosis
incidence

H0: there is no relationship between PM 2.5 exposure and tuberculosis
incidence

H0: there is no association between total green gas houses and
tuberculosis incidence

Ha: there is a relationship between proportion of urban population
living in slum areas and tuberculosis.

Ha: there is a relationship between safely managed sanitation and
tuberculosis incidence

Ha: there is a relationship between drinking water and tuberculosis
incidence

Ha: there is a relationship between fuel for cooking and tuberculosis
incidence

Ha: there is a relationship between PM 2.5 exposure and tuberculosis
incidence

Ha: there is a association between total green gas houses and
tuberculosis incidence

To cite other work (important everywhere, but likely happens first in
introduction), make sure your references are in the bibtex file
specified in the YAML header above and have the right bibtex key. Then
you can include like this:

Examples of reproducible research projects can for instance be found in
[@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning
processes, and the analysis approaches. You might want to provide a
shorter description here and all the details in the supplement.*

I haven't decide about data analysis methods. However, I am thinking to
perform time series analysis and multi linear regression (if suitable)
or Gene-realized linear model (we will see after conducting data
exploration). I am still in the process of reading more papers to get
more information which methods best suit this data and answer the
research questions.

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not
created with code (if you can do it with code, do it). @fig-schematic is
an example of some - completely random/unrelated - schematic that was
generated with Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition

Most of the data is downloaded from Worlbank Dataset. However, those
data are originally from many different courses/ organizations, such as
World Bank, WHA, UNICEF, UN-HABITAT, EDGAR, and IHME.

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for
analysis. Since this will be fairly long code for most dataset, it might
be a good idea to have it in one or several R scripts. If that is the
case, explain here briefly what kind of cleaning/processing you do, and
provide more details and well documented code somewhere (e.g. as
supplement in a paper). All materials, including files that contain
code, should be commented well so everyone can follow along.*

Dataset are either in excel or CSV files. Several steps from loading to
data analysis:

-   to load the data set in this part using read_excel or read_csv.

-   to combine them base on country name/ code and years.

<!-- -->

-   to drop unwanted variables

-   to select Asian and African Countries only

-   to drop N/A or missing value

-   to conduct data exploration

-   to decide suitable analysis method

-   to perform data analysis

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your
data. Show the most important descriptive results here. Additional ones
should go in the supplement. Even more can be in the R and Quarto files
that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the
`../../` notation. (Two dots means a folder up). You never want to
specify an **absolute** path like `C:\ahandel\myproject\results\`
because if you share this with someone, it won't work for them since
they don't have that path. You can also use the `here` R package to
create paths. See examples of that below. I generally recommend the
`here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could
compute simple statistics (e.g. simple models with 1 predictor) to look
for associations between your outcome(s) and each individual predictor
variable. Though note that unless you pre-specified the outcome and main
exposure, any "p\<0.05 means statistical significance" interpretation is
not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to
analyze your data and to produce meaningful figures, tables, etc. This
might again be code that is best placed in one or several separate R
scripts that need to be well documented. You want the code to produce
figures and data ready for display as tables, and save those. Then you
load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your
analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers
published using a fully reproducible setup similar to the one shown in
this template.

Note that this cited reference will show up at the end of the document,
the reference formatting is determined by the CSL file specified in the
YAML header. Many more style files for almost any journal [are
available](https://www.zotero.org/styles). You also specify the location
of your bibtex reference file in the YAML. You can call your reference
file anything you like.

{{< pagebreak >}}

# References