### 2025-02-23 code/eda-code/eda.qmd

---
title: "An example exploratory analysis script"
date: "2024-02-07"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I \# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
install.packages("maps")
library(maps)
install.packages("ggcorrplot")
library(ggcorrplot)
install.packages("plm")
library(plm)
install.packages("gt")
library(gt)
install.packages("car")
library(car)
install.packages("MASS")  # Jika belum terinstall
library(MASS)
install.packages("tidyverse")
library(tidyverse)
install.packages("reshape2")  # Install the package (only needed once)
library(reshape2)  # Load the package
install.packages("scales")
library(scales)

library(here)
library(knitr)
install.packages("kableExtra")
library(kableExtra)
install.packages("GGally")
library(GGally)
# Install and load the patchwork package
install.packages("patchwork")
library(patchwork)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","clean_noslum.rds")

#load data
mydata <- readRDS(data_location)

#subset for south east and Asia Africa only 
asia_africa <- mydata %>% 
  filter(region == c("South-East Asia", "Africa"))
unique(asia_africa$country)

saveRDS(asia_africa, here("data", "processed-data", "asia_africa.rds"))
```

# Data exploration through tables

```{r}

# Define distinct colors for each country
country_colors <- c(
  "Algeria" = "#E6194B",  # Red
  "Bangladesh" = "#3CB44B",  # Green
  "Central African Republic" = "#FFE119",  # Yellow
  "Chad" = "#4363D8",  # Blue
  "Ethiopia" = "#F58231",  # Orange
  "Ghana" = "#911EB4",  # Purple
  "Indonesia" = "#42D4F4",  # Light Blue
  "Lesotho" = "#F032E6",  # Pink
  "Madagascar" = "#FABEBE",  # Light Red
  "Malawi" = "#469990",  # Teal
  "Myanmar" = "#E6BEFF",  # Lavender
  "Nepal" = "#9A6324",  # Brown
  "Senegal" = "#FFFAC8",  # Cream
  "Sri Lanka" = "#800000",  # Maroon
  "Tanzania" = "#A9A9A9",  # Gray
  "Togo" = "#FFD8B1",  # Peach
  "Uganda" = "#000075",  # Dark Blue
  "Zimbabwe" = "#808000"  # Olive
)

# Convert country to a factor
asia_africa$country <- as.factor(asia_africa$country)

# Create boxplot with borders
plot1 <- ggplot(asia_africa, aes(x = country, y = tbi, fill = country)) +
  geom_boxplot(color = "black", size = 0.1) +  # Add black border with thicker lines
  scale_fill_manual(values = country_colors) +  # Manually assign colors
  labs(title = "Boxplot of TBI by Country",
       x = "Country",
       y = "TBI") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),  # Rotate labels for readability
        panel.border = element_rect(color = "black", fill = NA, size = 1.5))  # Add outer frame

plot_box <- here( "code","eda-code", "figures", "boxplot.png") # to set up location for the pictures created 
ggsave(filename = plot_box, plot=plot1, width = 12, height = 8, units = "in", dpi = 300) # save the pictures created

plot(plot1)

```

```{r}

# Select only numeric columns (excluding country, year, and region)
asia_africa_long <- asia_africa %>%
  select(-region) %>%  # Exclude categorical variable
  pivot_longer(cols = -c(country, year),  # Convert all numeric variables except country & year
               names_to = "Variable", 
               values_to = "Value")

# Create line plots for all variables
plot2 <- ggplot(asia_africa_long, aes(x = year, y = Value, group = country, color = country)) +
  geom_line(size = 1, alpha = 0.8) +  # Alpha for transparency
  facet_wrap(~ Variable, scales = "free_y") +  # Separate plots for each variable
  labs(title = "Trends of All Variables Over Time",
       x = "Year",
       y = "Value",
       color = "Country") +
  theme_minimal() +
  theme(legend.position = "none",  # Hide legend for clarity
        axis.text.x = element_text(angle = 45, hjust = 1))

plot_spageti <- here( "code","eda-code", "figures", "spageti.png") # to set up location for the pictures created 
ggsave(filename = plot_spageti, plot=plot2, width = 16, height = 12, units = "in", dpi = 300) # save the pictures created

plot(plot2)
```

```{r}
plot3 <- ggplot(asia_africa, aes(x = year, y = tbi, group = country, color = country)) +
  geom_line(alpha = 0.7) +  # Transparency to see overlapping lines
  labs(title = "Spaghetti Plot of TBC Incidence",
       x = "Year",
       y = "TBC Incidence") +
  theme_minimal() +
  theme(legend.position = "none")

plot_spageti_all <- here( "code","eda-code", "figures", "spageti-tbi.png") # to set up location for the pictures created 
ggsave(filename = plot_spageti_all, plot=plot3, width = 16, height = 12, units = "in", dpi = 300) # save the pictures created

plot(plot3)
```

***Creating plots to show the trend of all variables in each country over the year***

```{r}
# Find a scaling factor to adjust TBI to match percentage values
max_indicator <- max(asia_africa[, c("pro_safesanitation", "pro_basicsanitation", "pro_safewater", "pro_basicwater", "pro_cleanfuel")], na.rm = TRUE)
max_tbi <- max(asia_africa$tbi, na.rm = TRUE)
scale_factor <- max_indicator / max_tbi  # Scale TBI to fit

# Reshape the data from wide to long format
asia_africa_long <- asia_africa %>%
  pivot_longer(cols = c(pro_safesanitation, pro_basicsanitation, pro_safewater, pro_basicwater, pro_cleanfuel),
               names_to = "Indicator", 
               values_to = "Value")

# Create the plot
plot4 <- ggplot() +
  # First, plot water, sanitation, and clean fuel indicators
  geom_line(data = asia_africa_long, 
            aes(x = year, y = Value, color = Indicator, group = Indicator), 
            size = 1) +
  # Add TBI as a separate line with a different scale
  geom_line(data = asia_africa, 
            aes(x = year, y = tbi * scale_factor, color = "TBI"), 
            size = 1, linetype = "dashed") + 
  facet_wrap(~ country) +  # Separate plots for each country
  theme_minimal() +
  labs(title = "Percentage of Population with Access to Water, Sanitation, Clean Fuel & TBI Trends",
       x = "Year",
       y = "Percentage (%)",
       color = "Indicator") +  # Label for legend
  scale_y_continuous(sec.axis = sec_axis(~ . / scale_factor, name = "TBI")) +  # Secondary Y-axis for TBI
  scale_color_manual(values = c("blue", "green", "purple", "orange", "brown", "red")) +  # Custom colors
  theme(legend.position = "bottom",  
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.border = element_rect(color = "black", fill = NA, size = 1))  # Rotate x-axis labels for readability

plot_location1 <- here( "code","eda-code", "figures", "line-graph-tbivsfactors.png") # to set up location for the pictures created 
ggsave(filename = plot_location1, plot=plot4, width = 12, height = 16, units = "in", dpi = 300) # save the pictures created 
plot(plot4)
```

Check the layout after saving the figures

```{r}

#| label: fig-multivariable 
#| fig-cap: "Line charts of all variables"
#| echo: FALSE
knitr::include_graphics(here("code", "eda-code", "figures", "line-graph-tbivsfactors.png"))
```

***I want to show the mean of each variable every year based on regions***

```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(kableExtra)
library(here)

# Define row colors for each category
row_colors <- c(
  "TBI" = "#FFD700", # Gold
  "GDP" = "#ADD8E6", # Light Blue
  "HDI" = "#90EE90", # Light Green
  "Clean Fuel" = "#FFB6C1", # Light Pink
  "Safe Water" = "#87CEEB", # Sky Blue
  "Basic Sanitation" = "#FFA07A", # Light Salmon
  "Safe Sanitation" = "#F08080", # Light Coral
  "PM2.5" = "#D3D3D3", # Light Gray
  "Greenhouse" = "#B0C4DE" # Light Steel Blue
)

# Create summary per year
summary_table <- asia_africa %>%
  group_by(year) %>%  
  summarise(
    Min_TBI = min(tbi, na.rm = TRUE),
    Mean_TBI = mean(tbi, na.rm = TRUE),
    Max_TBI = max(tbi, na.rm = TRUE),
    SD_TBI = sd(tbi, na.rm = TRUE),

    Min_GDP = min(gdp, na.rm = TRUE),
    Mean_GDP = mean(gdp, na.rm = TRUE),
    Max_GDP = max(gdp, na.rm = TRUE),
    SD_GDP = sd(gdp, na.rm = TRUE),

    Min_HDI = min(hdi, na.rm = TRUE),
    Mean_HDI = mean(hdi, na.rm = TRUE),
    Max_HDI = max(hdi, na.rm = TRUE),
    SD_HDI = sd(hdi, na.rm = TRUE),

    Min_CleanFuel = min(pro_cleanfuel, na.rm = TRUE),
    Mean_CleanFuel = mean(pro_cleanfuel, na.rm = TRUE),
    Max_CleanFuel = max(pro_cleanfuel, na.rm = TRUE),
    SD_CleanFuel = sd(pro_cleanfuel, na.rm = TRUE),

    Min_SafeWater = min(pro_safewater, na.rm = TRUE),
    Mean_SafeWater = mean(pro_safewater, na.rm = TRUE),
    Max_SafeWater = max(pro_safewater, na.rm = TRUE),
    SD_SafeWater = sd(pro_safewater, na.rm = TRUE),

    Min_BasicSanitation = min(pro_basicsanitation, na.rm = TRUE),
    Mean_BasicSanitation = mean(pro_basicsanitation, na.rm = TRUE),
    Max_BasicSanitation = max(pro_basicsanitation, na.rm = TRUE),
    SD_BasicSanitation = sd(pro_basicsanitation, na.rm = TRUE),

    Min_SafeSanitation = min(pro_safesanitation, na.rm = TRUE),
    Mean_SafeSanitation = mean(pro_safesanitation, na.rm = TRUE),
    Max_SafeSanitation = max(pro_safesanitation, na.rm = TRUE),
    SD_SafeSanitation = sd(pro_safesanitation, na.rm = TRUE),

    Min_PM2.5 = min(mean_exposure_pm2.5, na.rm = TRUE),
    Mean_PM2.5 = mean(mean_exposure_pm2.5, na.rm = TRUE),
    Max_PM2.5 = max(mean_exposure_pm2.5, na.rm = TRUE),
    SD_PM2.5 = sd(mean_exposure_pm2.5, na.rm = TRUE),

    Min_Greenhouse = min(total_greenhouses, na.rm = TRUE),
    Mean_Greenhouse = mean(total_greenhouses, na.rm = TRUE),
    Max_Greenhouse = max(total_greenhouses, na.rm = TRUE),
    SD_Greenhouse = sd(total_greenhouses, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -year, names_to = "Statistic", values_to = "Value") %>%
  pivot_wider(names_from = year, values_from = Value) %>% # Convert to wide format
  rowwise() %>% 
  mutate(Accumulated = mean(c_across(where(is.numeric)), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(across(where(is.numeric), ~ sprintf("%.2f", .x))) %>% # Format numbers to 2 decimal places
  relocate(Accumulated, .after = Statistic) %>% # Move Accumulated after Statistic
  mutate(Row_Index = row_number(), # Create row indices for coloring
         Category = case_when(
           grepl("TBI", Statistic) ~ "TBI",
           grepl("GDP", Statistic) ~ "GDP",
           grepl("HDI", Statistic) ~ "HDI",
           grepl("CleanFuel", Statistic) ~ "Clean Fuel",
           grepl("SafeWater", Statistic) ~ "Safe Water",
           grepl("BasicSanitation", Statistic) ~ "Basic Sanitation",
           grepl("SafeSanitation", Statistic) ~ "Safe Sanitation",
           grepl("PM2.5", Statistic) ~ "PM2.5",
           grepl("Greenhouse", Statistic) ~ "Greenhouse"
         ))

# Generate table with colors
summary_table_display <- summary_table %>%
  select(-Category, -Row_Index) %>% # Remove Category before displaying
  kable(caption = "Summary Statistics by Year with Accumulated Column") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# Apply row coloring dynamically
for (cat in names(row_colors)) {
  rows_to_color <- which(summary_table$Category == cat) # Get row indices
  summary_table_display <- summary_table_display %>%
    row_spec(rows_to_color, background = row_colors[cat])
}

saveRDS(summary_table_display, file = here("code", "eda-code", "tables", "summary_table.html"))
print(summary_table_display)

```

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Summary Statistics Table"
#| echo: FALSE

htmltools::includeHTML(here("code", "eda-code", "tables", "summary_table.html"))

```

## Data Analysis

***Test the linearity***

```{r}
gg1 <- ggplot(asia_africa, aes(x = gdp, y = tbi)) +
  geom_point() +  # Create scatter plot
  ggtitle("Scatter Plot: tbi vs GDP") +  # Add title
  xlab("GDP") +  # Label the x-axis
  ylab("TBI") +  # Label the y-axis
  theme_minimal()

gg2 <- 
```

```{r}

# Create a list of scatter plots for each independent variable against tbi
plots <- lapply(independent_vars, function(var) {
  ggplot(asia_africa, aes_string(x = var, y = "tbi")) +
    geom_point() + 
    ggtitle(paste("tbi vs", var)) + 
    xlab(var) + 
    ylab("TBI") + 
    theme_minimal()
})

# Combine all plots into a single frame
combined_plot <- wrap_plots(plots, ncol = 3)  # Adjust ncol to the number of columns you want
combined_plot

combineplot_location <- here( "code","eda-code", "figures", "scatterplot.png") # to set up location for the pictures created 
ggsave(filename = combineplot_location, plot=combined_plot, width = 12, height = 12, units = "in", dpi = 300) # save the pictures created 
plot(combined_plot)

```

 Spearman's Rank Correlation (Recommended) Spearman correlation is a non-parametric method that measures monotonic relationships (whether the variables move together, not necessarily linearly).

```{r}

# Compute Spearman correlation matrix
cor_matrix_spearman <- cor(asia_africa[, c("tbi","gdp", "hdi", "pro_cleanfuel", "pro_safewater", 
                                           "pro_basicsanitation", "pro_safesanitation", "mean_exposure_pm2.5", "total_greenhouses")], method = "spearman", use = "complete.obs")

# Melt the matrix into long format for ggplot

cor_melted <- melt(cor_matrix_spearman)

# Plot heatmap
plot5 <- ggplot(cor_melted, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
  geom_text(aes(label = round(value, 2)), size = 4) +  # Show correlation values
  theme_minimal() +
  labs(title = "Spearman Correlation Heatmap", x = "", y = "")
matrix_location <- here( "code","eda-code", "figures", "correlation_matrix.png") # to set up location for the pictures created 
ggsave(filename = matrix_location, plot=plot5, width = 12, height = 12, units = "in", dpi = 300) # save the pictures created 
plot(plot5)
```

```{r}

#| label: fig-Spearman's Correlation Matrix 
#| fig-cap: "Spearman's Correlation Matrix "
#| echo: FALSE
knitr::include_graphics(here("code", "eda-code", "figures", "correlation_matrix.png"))
```

I will performed longitudinal data analysis

```{r}
library(car)
#vif(lm(tbi ~ gdp + hdi + pro_cleanfuel + pro_safewater + pro_basicsanitation + 
       pro_safesanitation + mean_exposure_pm2.5 + total_greenhouses, data = pdata)) # The error might be occure due to multicolinearity

#alias(lm(tbi ~ gdp + hdi + pro_cleanfuel + pro_safewater + pro_basicsanitation + 
         pro_safesanitation + mean_exposure_pm2.5 + total_greenhouses, data = asia_africa))

stepwise_model <- stepAIC(lm(tbi ~ gdp + hdi + pro_cleanfuel + pro_safewater + pro_basicsanitation + 
                             pro_safesanitation + mean_exposure_pm2.5 + total_greenhouses, 
                             data = asia_africa), direction = "both") # drop variable with multicolinearity

summary(stepwise_model)

```

***Performing longitudinal analysis***

Test for mixed or fixed effect using Hausman Test

```{r}
pdata <- pdata.frame(asia_africa, index = c("country", "year"))

fe_model <- plm(tbi ~ pro_basicsanitation + mean_exposure_pm2.5, 
                data = pdata, model = "within")
re_model <- plm(tbi ~ pro_basicsanitation + mean_exposure_pm2.5, 
                data = pdata, model = "random")

hausman_test <- phtest(fe_model, re_model)
print(hausman_test)

```

Based on the Hausman Test, fixed effect is not consistent. Therefore, I will use mixed effect test.

```{r}
summary(re_model)
```

Interpretation: ***1. variance components Analyisis***

88% dari total variasi berasal dari perbedaan antar negara, menunjukkan bahwa perbedaan karakteristik tiap negara sangat berpengaruh terhadap insiden TBC.

12% berasal dari variasi idiosinkratik, atau faktor yang berubah dari waktu ke waktu dalam suatu negara.

***2. Theta***

Rata-rata theta = 0.8868, menunjukkan bahwa sebagian besar variasi dalam data disebabkan oleh perbedaan antar negara. Theta yang mendekati 1 berarti bahwa individu (negara) memiliki efek spesifik yang kuat dibandingkan variasi waktu.