### 2025-02-23 products/manuscript/Manuscript.qmd

---
title: "The effects of multi environmental exposures on Tuberculosis Incidence in South East Asian and African countris"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

This uses MS Word as output format. [See
here](https://quarto.org/docs/output-formats/ms-word.html) for more
information. You can switch to other formats, like html or pdf. See [the
Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

-   Muhammad Nasir$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
-   Second Author$^{3, *}$
-   Third Author$^{3, \dagger}$
-   Last Author$^{1, 2, \land}$

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.
2.  Faculty of Public Health, Universitas Tadulako, Palu, Indonesia.
3.  Yet another Affiliation.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: some\@email.com

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

Tuberculosis is old infectious disease which still exists and become
major public health in the world, particularly in developing countries.
Tuberculosis is linked to environmental factors. This study will analyze
the effects of multi environmental factors on tuberculosis incidence in
Asia and Africa regions, where the majority of tuberculosis coming from
these regions. The environmental factors and country economic and human
development index as covariates.

## Description of data and data source

The study will use multi sources of data including World Bank, World
Health Organization (WHO), United Nation for Children Fund (Unicef),
UN-HABITAT, Emissions Database for Global Atmospheric Research (EDGAR),
and Institute for Health Metrics and Evaluation (IHME). Different
dataset contains different variable, however all dataset have same year,
country name and country codes that can be used to combine the dataset.
The data also contain information of countries globally from 2000 to
2020. The variables will include proportion of urban population living
in slum areas, percentage of population using safely managed sanitation
services, percentage of population using at least basic drinking water
service, fuel for cooking, Population-weighted exposure to ambient PM2.5
pollution, total green houses, and will include GDP/ country economic
status, human development index, and Universal Health Coverage as
covariates.

Note: I have downloaded the dataset, but I have not conducted any
cleaning or exploration on the dataset. List of data sources and website
is provided in an excel fine in my data folder. I will include them in
my reference for the next steps.

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

The main aim of this study is to look at the impacts of environmental
factors on tuberculosis incidence.

RQ: What environmental factors associated with tuberculosis incidence?

H0: there is no relationship between proportion of urban population
living in slum areas and tuberculosis.

H0: thee is no relationship between safely managed sanitation and
tuberculosis incidence

H0: there is no relationship between drinking water and tuberculosis
incidence

H0: there is no relationship between fuel for cooking and tuberculosis
incidence

H0: there is no relationship between PM 2.5 exposure and tuberculosis
incidence

H0: there is no association between total green gas houses and
tuberculosis incidence

Ha: there is a relationship between proportion of urban population
living in slum areas and tuberculosis.

Ha: there is a relationship between safely managed sanitation and
tuberculosis incidence

Ha: there is a relationship between drinking water and tuberculosis
incidence

Ha: there is a relationship between fuel for cooking and tuberculosis
incidence

Ha: there is a relationship between PM 2.5 exposure and tuberculosis
incidence

Ha: there is a association between total green gas houses and
tuberculosis incidence

To cite other work (important everywhere, but likely happens first in
introduction), make sure your references are in the bibtex file
specified in the YAML header above and have the right bibtex key. Then
you can include like this:

Examples of reproducible research projects can for instance be found in
[@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

The datasets were loded one by one, then drop all unused variable before
comined those datasets using reduce() function. Countries, which have at
least one missing years from 2000 to 2020, were dropped from the study.
To make sure that the data is clean and ready to analyse, all missing
values were dropped. This study focuses on South East Asia and Africa
Regions, where the major proportion of tuberculosis cases came.
Therefore, a newsubset for South East Asia and African countries was
created.

After data cleaning, multiple exploratory data analysis was forformed
such us summary and vizualising the data. A bloxplot was created to show
Tuberculosis Incidence in each country during the time frame. Spagheti
plot was created to visualize the trend of tuberculosis incidence and
other variables over the time. Lastly, line charts showing the trend of
all variables in each country.

The exploratory data analysis showed that the data were not normally
distributed and there was no linear relationship between tuberculosis
incidence and other predictors. Therefore, Spearman's correlation matrix
was perfomed to see the correlations. To have better understanding, a
heatmap matrix plot was created to show the correlation.

The main statistical analysis in thie study is longitudinal data
analysis (eithre fixed effect or mixed effect method). To see which one
is better, Hausman Test was performed to determine which model is
better.

NOTE: I still need to discuss and need more advice about the data
analysis from Dr. Handel

## Schematic of workflow

NOTE: I AM STILL WORKING TO FIGUR THIS ONE OUT Sometimes you might want
to show a schematic diagram/figure that was not created with code (if
you can do it with code, do it). @fig-schematic is an example of some -
completely random/unrelated - schematic that was generated with
Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition

Most of the data is downloaded from Worlbank Dataset. However, those
data are originally from many different courses/ organizations, such as
World Bank, WHA, UNICEF, UN-HABITAT, EDGAR, and IHME.

## Data import and cleaning

Data Import and data cleaning process have been explained in the methode
part. To see the code for data import and data cleaning , please visit
working directory ("code", "processing-code", "processingfile.qdm").
Processed and clean data are saved at a folder ("data",
"processed-data"). There are many .rds files in the folder icnlduing
file for each variables, and file after combined all variables.

Dataset are either in excel or CSV files. Several steps from loading to
data analysis:

-   to load the data set in this part using read_excel or read_csv.

-   to combine them base on country name/ code and years.

-   to drop unwanted variables

-   to select Asian and African Countries only

-   to drop N/A or missing value

-   to conduct data exploration

-   to decide suitable analysis method

-   to perform data analysis

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

There are some exploratory data analysis conducted and resulted table,
and charts. Table 1 shows the summary statistics including min, mean,
max, and standard deviation of all variables.

NOTE: I still have issue in saving the stiled tables. I am still working
on it. I will show it next step.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

The line graph ( @fig-1 ) shows the trend of all variables over the
period in each country. It can be seen that the majority of country
showed the decreases trend in tuberculosis incidence over 21 years
period, except Lesotho showing a sharp increase until the middle of the
period before dramatic decrease until the end of the period. All
countries showed the increase trend in percentage of population using
using basic and safe drinking water, basic and safe sanitation, and
clean fuel and technology for cooking, except Zimbabuwe showing the
decrease in percentage of population using basic drinking water and safe
sanitation, and Senegal showing the decrease of population using clean
fuel and technology for cooking.

```{r}

#| label: fig-1
#| fig-cap: "Trends of all variable over the period in each country"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","line-graph-tbivsfactors.png"))

```

( @fig-2 ) shows spaghetti plots in each variables. We can see the
trends of each country in every variables. It can be seen that generally
most of the countries show the increase trends in water and sanitation,
clean fuel use, human development index, total greenhouses, but
fluctuated trends in mean of PM2.5 expousures.

```{r}

#| label: fig-2
#| fig-cap: "Spaghetti plots in each variables"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","spageti.png"))

```

Now lets take a more detail tuberculosis incidence in each country over
the year using the following box plot. Lesotho leads the incidence
followed by Myanmar and Tanzania.

```{r}

#| label: fig-3
#| fig-cap: "Spaghetti plots in each variables"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","boxplot.png"))

```

## Basic statistical analysis

As first step of statistical analysis. Using scatterplot to show the linearity between tuberculosis incidence and predictors. ( @fig-4 ) shows the scatterplots. The scatterplots do not show linear relationship between tuberculosis Incidence and predictors.

```{r}

#| label: fig-4
#| fig-cap: "Scatter plots between Tuberculosis Incidence and Predictors "
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","scatterplot.png"))

```
Before going further, we need to look at correlation between tuberculosis Incidence and all predictors.

Since they show nonlinearity, spearman's correlation performed and hsowing in the following correlation matrix. See ( @fig-5 )

```{r}

#| label: fig-5
#| fig-cap: "Spearman Correlation Matrix "
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","correlation_matrix.png"))

```

Since there is no linear relationship, nonlinear relationship test was performed.

```{r}

```

*To get some further insight into your data, if reasonable you could
compute simple statistics (e.g. simple models with 1 predictor) to look
for associations between your outcome(s) and each individual predictor
variable. Though note that unless you pre-specified the outcome and main
exposure, any "p\<0.05 means statistical significance" interpretation is
not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to
analyze your data and to produce meaningful figures, tables, etc. This
might again be code that is best placed in one or several separate R
scripts that need to be well documented. You want the code to produce
figures and data ready for display as tables, and save those. Then you
load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your
analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers
published using a fully reproducible setup similar to the one shown in
this template.

Note that this cited reference will show up at the end of the document,
the reference formatting is determined by the CSL file specified in the
YAML header. Many more style files for almost any journal [are
available](https://www.zotero.org/styles). You also specify the location
of your bibtex reference file in the YAML. You can call your reference
file anything you like.

{{< pagebreak >}}

# References