### 2025-03-19 products/manuscript/Manuscript.qmd

---
title: "The effects of Water and Sanitation on Tuberculosis Incidence in South East Asian and African countris"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

This uses MS Word as output format. [See
here](https://quarto.org/docs/output-formats/ms-word.html) for more
information. You can switch to other formats, like html or pdf. See [the
Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

-   Muhammad Nasir$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
-   Second Author$^{3, *}$
-   Third Author$^{3, \dagger}$
-   Last Author$^{1, 2, \land}$

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.
2.  Faculty of Public Health, Universitas Tadulako, Palu, Indonesia.
3.  Yet another Affiliation.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: some\@email.com

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

Tuberculosis is old infectious disease which still exists and become
major public health in the world, particularly in developing countries.
Tuberculosis is linked to environmental factors. This study will analyze
the effects of multi environmental factors, specially water and
sanitation on tuberculosis incidence in Asia and Africa regions, where
the majority of tuberculosis coming from these regions. The water and
sanitation will be the main exposures and Human Development Index as a
control variable.

## Description of data and data source

The study will use multi sources of data including World Bank, World
Health Organization (WHO), United Nation for Children Fund (Unicef).
Different dataset contains different variable, however all dataset have
same year, country name and country codes that can be used to combine
the dataset. The data also contain information of countries globally
from 2000 to 2020. The variables will include percentage of population
using safely managed sanitation services, percentage of population using
at least basic drinking water service, percentage of population using
well-managed water, percentage of population using basic sanitation,
percentage of population using well-managed sanitation, and human
development index.

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

The main aim of this study is to look at the effect of Water and
Sanitation on on tuberculosis incidence.

RQ: Does access to water and sanitation effect Tuberculosis incidence in
Asian and African Regions?

Null Hypotheisis

H0: thee is no effect of access to basic sanitation on tuberculosis
incidence

H0: thee is no effect of access to safely managed sanitation and on
tuberculosis incidence

H0: there is no effect of access to access basic drinking water and
tuberculosis incidence

H0: there is no effect of access to safely managed drinking water on
tuberculosis incidence

Alternative Hypothesis Ha: thee is an effect of access to basic
sanitation on tuberculosis incidence

Ha: thee is an effect of access to safely managed sanitation and on
tuberculosis incidence

Ha: there is an effect of access to access basic drinking water and
tuberculosis incidence

Ha: there is an effect of access to safely managed drinking water on
tuberculosis incidence

To cite other work (important everywhere, but likely happens first in
introduction), make sure your references are in the bibtex file
specified in the YAML header above and have the right bibtex key. Then
you can include like this:

Examples of reproducible research projects can for instance be found in
[@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

The datasets were loaded one by one, then drop all unused variable
before combined those datasets using reduce() function. Countries, which
have at least one missing years from 2000 to 2020, were dropped from the
study. To make sure that the data is clean and ready to analyse, all
missing values were dropped. This study focuses on South East Asia and
Africa Regions, where the major proportion of tuberculosis cases came.
Therefore, a newsubset for South East Asia and African countries was
created.

After data cleaning, multiple exploratory data analysis was performed
such us summary and visualizing the data. A bloxplot was created to show
Tuberculosis Incidence in each country during the time frame. Spaghetti
plot was created to visualize the trend of tuberculosis incidence and
other variables over the time. Lastly, line charts showing the trend of
all variables in each country.

The exploratory data analysis showed that the data were not normally
distributed and there was no linear relationship between tuberculosis
incidence and other predictors. Therefore, Spearman's correlation matrix
was performed to see the correlations. To have better understanding, a
heat map matrix plot was created to show the correlation.

The main statistical analysis in the study is longitudinal data analysis
(either fixed effect or mixed effect method). To see which one is
better, Hausman Test was performed to determine which model is better.

NOTE: I still need to discuss and need more advice about the data
analysis from Dr. Handel

## Schematic of workflow

NOTE: I AM STILL WORKING TO FIGUR THIS ONE OUT Sometimes you might want
to show a schematic diagram/figure that was not created with code (if
you can do it with code, do it). @fig-schematic is an example of some -
completely random/unrelated - schematic that was generated with
Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data acquisition

The data was downloaded from Worlbank Dataset. However, those data are
originally from many different courses/ organizations, such as World
Bank, WHA, UNICEF, UN-HABITAT, EDGAR, and IHME.

## Data import and cleaning

Data Import and data cleaning process have been explained in the method
part. To see the code for data import and data cleaning , please visit
working directory ("code", "processing-code", "processingfile.qdm").
Processed and clean data are saved at a folder ("data",
"processed-data"). There are many .rds files in the folder indulging
file for each variables, and file after combined all variables.

Dataset are either in excel or CSV files. Several steps from loading to
data analysis:

-   to load the data set in this part using read_excel or read_csv.

-   to combine them base on country name/ code and years.

-   to drop unwanted variables

-   to select Asian and African Countries only

-   to drop N/A or missing value

-   to conduct data exploration

-   to decide suitable analysis method

-   to perform data analysis

## Statistical analysis

*Explain anything related to your statistical analyses.*

The study applied Generalized Additive Mixed Model (GAMM). The model can
capture mixed effect of countries on Tuberculosis Incidence with non
linear relationship. The data analysis started by looking at the
relationship between Tuberculosis Incidence and Predictors. Non linear
relationship was discovered.Next step is to perform non-linear
relationship to looking at form of relationship. Human Development Index
(HDI), Access to basic water and well-managed water show curvature
relationship, while access to basic sanitation and well-managed
sanitation show polynomial. Data transformation was performed based.
After data transformation, we fit model1, which only includes water and
sanitation variables (as main variables in this study), model2 includes
all predictors (HDI, access to water and sanitation), model3 only
includes significant variables in the previous model. Model 1, 2 and 3
use Generalized Linear Mixed Model (GLMM). We also performed VIF test to
look at multicollinearity. Model 4 uses Generalized Additive Mixed
Models. Moreover, BIC and AIC tests were performed to look at the best
model. BIC and AIC tests found that Model 4 (GAMM) performs better with
the smallest BIC and AIC. To be able to interpret GAMM better, we
visualize the result into plots.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

There are some exploratory data analysis conducted and resulted table,
and charts. Table 1 shows the summary statistics including min, mean,
max, and standard deviation of all variables.

NOTE: I still have issue in saving the tables. I am still working on it.
I will show it next step.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS (here("code", "eda-code", "tables", "summary_table.rds"))
knitr::kable(resulttable)
```

The line graph ( @fig-1 ) shows the trend of all variables over the
period in each country. It can be seen that the majority of country
showed the decreases trend in tuberculosis incidence over 21 years
period. All countries showed the increase trend in percentage of
population using using basic and safe drinking water, basic and safe
sanitation, except Zimbabuwe showing the decrease in percentage of
population using basic drinking water and safe sanitation.

```{r}

#| label: fig-1
#| fig-cap: "Trends of all variable over the period in each country"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","line-graph-tbivsfactors.png"))

```

( @fig-2 ) shows spaghetti plots in each variables. We can see the
trends of each country in every variables. It can be seen that generally
most of the countries show the increase trends in water and sanitation,
and human development index

```{r}

#| label: fig-2
#| fig-cap: "Spaghetti plots in each variables"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","spageti.png"))

```

Now lets take a more detail tuberculosis incidence in each country over
the year using the following box plot. Lesotho leads the incidence
followed by Myanmar and Tanzania.

```{r}

#| label: fig-3
#| fig-cap: "Boxplot plots for each Country"
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","boxplot.png"))

```

## Basic statistical analysis

As first step of statistical analysis. Using scatter plot to show the
linearity between tuberculosis incidence and predictors. ( @fig-4 )
shows the scatter plots. The scatter plots do not show linear
relationship between tuberculosis Incidence and predictors.

```{r}

#| label: fig-4
#| fig-cap: "Scatter plots between Tuberculosis Incidence and Predictors "
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","scatterplot.png"))

```

Before going further, we need to look at correlation between
tuberculosis Incidence and all predictors.

Since they show non linearity, spearman's correlation performed and
showing in the following correlation matrix. See ( @fig-5 )

```{r}

#| label: fig-5
#| fig-cap: "Spearman Correlation Matrix "
#| echo: FALSE
knitr::include_graphics(here("code","eda-code","figures","correlation_matrix.png"))

```

Now, we want to assess non linear relationship.

```{r}

#| label: fig-6
#| fig-cap: "Nonlinear relationship"
#| echo: FALSE
knitr::include_graphics(here("code","analysis-code","figures","nonlinear.png"))

```

After performing nonlinear model, data transformation used based on the
shave or distribution of the plots from nonlinear relationship plots (
@fig-6 ) above. ( @fig-7 ) shows the data distribution after data
transformation.

```{r}

#| label: fig-7
#| fig-cap: "transformed data plots"
#| echo: FALSE
knitr::include_graphics(here("code","analysis-code","figures","transformed.png"))

```

## Full analysis

There are four models build; fitting model with only water and
sanitation variables, then fitting models including co-variate (HDI).
The third model is to fit only significant variables. The first three
models use Generalized Linear Mixed Models (GLMM), while the last model
used Generalized Additive Mixed Models (GAMM). Those models were
compared using AIC and BIC criteria to select the best/ the final
models.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Model Criteria"
#| echo: FALSE
resulttable2=readRDS (here("code", "analysis-code", "tables", "criteria_output.rds"))
knitr::kable(resulttable2)
```

Based on BIC and AIC criteria, the best model fit the data is
Generalized Additive Mixed Models.

\$ \log(\text{TBI}) \sim s(\text{hdi}) + s(\text{pro\_basicwater}) +
s(\text{pro\_safewater}) + s(\text{pro\_basicsanitation}) +
s(\text{country}, \text{bs}="re") \$

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Generalized Additive Mixed Models"
#| echo: FALSE
resulttable3=readRDS (here("code", "analysis-code", "tables", "summary_Gamm.rds"))
knitr::kable(resulttable3)
```

This study employed a Generalized Additive Model (GAM) with a Gaussian
family and identity link function to examine the association between
tuberculosis incidence and key environmental and socioeconomic factors
across multiple countries. The model incorporated smooth functions for
continuous predictors and a random effect for country to account for
within-country variability. The results indicate that HDI and basic
sanitation are significant determinants of tuberculosis incidence, while
access to basic and safe water did not show strong associations.

The smooth function for Human Development Index (HDI) was highly
significant (p \< 2e-16), with an estimated effective degrees of freedom
(EDF) = 4.283, confirming a non-linear relationship between HDI and
tuberculosis incidence. The effect was strongest at lower HDI values,
suggesting that improvements in socioeconomic conditions may have the
greatest impact in reducing tuberculosis burden in low-HDI countries.

Among the water and sanitation variables, basic water access was not
statistically significant (p = 0.73379), indicating no strong evidence
of an association between water coverage and tuberculosis incidence.
Similarly, safe water access was not significant at α = 0.05 (p =
0.05605), suggesting only a weak relationship. However, basic sanitation
access was statistically significant (p = 0.00542, EDF = 3.845),
suggesting that higher access to sanitation is associated with a
reduction in tuberculosis incidence. This highlights the importance of
improving sanitation infrastructure as a potential intervention for
tuberculosis control.

The random effect of country (s(country)) was highly significant (p \<
2e-16, EDF = 16.768), indicating substantial between-country variation
in tuberculosis incidence. This justifies the inclusion of random
effects to account for country-level heterogeneity and underscores the
need for context-specific policy interventions tailored to national
conditions.

The model fit was strong, with an adjusted R² of 0.974, indicating that
97.4% of the variance in tuberculosis incidence was explained by the
model. The scale estimate was 0.014598, suggesting a low residual
variance, which reflects a well-fitting model.

```{r}
#| label: fig-7
#| fig-cap: "Generalized Additive Mixed Models"
#| echo: FALSE
knitr::include_graphics(here("code","analysis-code","figures","gam_diagnostics.jpg"))

```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your
analysis.*

## Conclusions

*What are the main take-home messages?*

These findings underscore the significant role of socioeconomic
conditions and sanitation infrastructure in tuberculosis control. While
access to basic and safe water was not a significant predictor,
improvements in sanitation and overall socioeconomic development (HDI)
were strongly associated with lower tuberculosis incidence. Given the
significant country-level differences, tuberculosis prevention
strategies should be tailored to national contexts, emphasizing
improvements in living standards and sanitation access. These results
provide valuable insights for policymakers aiming to design targeted
public health interventions in high-burden settings.

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers
published using a fully reproducible setup similar to the one shown in
this template.

Note that this cited reference will show up at the end of the document,
the reference formatting is determined by the CSL file specified in the
YAML header. Many more style files for almost any journal [are
available](https://www.zotero.org/styles). You also specify the location
of your bibtex reference file in the YAML. You can call your reference
file anything you like.

{{< pagebreak >}}

# References

I will include reference for the sourse of data in the next step.