### 2025-04-08 assets/statistical-analysis.qmd
Edit manuscript

---
title: "statistical-analysis"
format: html
editor: visual
---

##

```{r}
#load needed packages. make sure they are installed.
library(here)
library(dplyr)
library(skimr)
library(ggplot2)
#install.packages("maps")
library(maps)
#install.packages("ggcorrplot")
library(ggcorrplot)
#install.packages("plm")
library(plm)
#install.packages("gt")
library(gt)
#install.packages("car")
library(car)
#install.packages("MASS")
library(MASS)
install.packages("tidyverse")
library(tidyverse)
#install.packages("reshape2") 
library(reshape2) 
#install.packages("scales")
library(scales)
library(splines)
library(here)
library(knitr)
library(kableExtra)
#install.packages("GGally")
library(GGally)
#install.packages("patchwork")
library(patchwork)
#install.packages("gridExtra")
library(gridExtra)
#install.packages("lmerTest")
library(lmerTest)
#install.packages("stargazer")
library(stargazer)
#install.packages("lmtest")
library(lmtest)
#install.packages("mgcv") 
library(mgcv)  
install.packages("gamm4")
library(gamm4)
install.packages("glmnet")
library(glmnet)
install.packages("brms")
library(brms)
```

Load the dataset

```{r}

seaadf_location <- here("data", "processed-data", "seaa.rds")
seaa_df <- readRDS(seaadf_location)

str(seaa_df)

seaa_df <- seaa_df %>%
  mutate(across(c(pro_basicwater, pro_safewater, 
                  pro_basicsanitation, pro_safesanitation), 
                ~ as.numeric(.)))
```

## Data Analysis

First, we perform linear regression. Due to high multicolinearity, percentage of population using safe sanitation is eliminated from the model. 
```{r}
# Fit a linear model (without Interaction)
model_lm1 <- lm(tbi ~ pro_basicwater + pro_safewater + 
               pro_basicsanitation +  hdi, 
               data = seaa_df)
summary(model_lm1)
```
```{r}
# Fit a linear model (with Interaction)
model_lm_inte <- lm(tbi ~ (pro_basicwater + pro_safewater + 
               pro_basicsanitation +  hdi)^2, 
               data = seaa_df)
summary(model_lm_inte )
```

```{r}
# Plot residuals vs. fitted values for the linear model
plot(model_lm1$fitted.values, resid(model_lm1), 
     xlab = "Fitted Values", ylab = "Residuals", 
     main = "Residuals vs. Fitted Plot")
abline(h = 0, col = "red")  # Horizontal line at 0
```

# Fit a non-linear model (GAM)
```{r}

model_gam1 <- gam(tbi ~ s(pro_basicwater) + s(pro_safewater) + s(pro_basicsanitation) + s(hdi), 
                 data = seaa_df, method = "REML")

```

compared the linear model and non linear model using AIC

```{r}

```

```{r}
model_gam_inte <- gam(tbi ~ 
                    s(pro_basicwater) + 
                    s(pro_safewater) + 
                    s(pro_basicsanitation) + 
                    s(hdi) + 
                    te(pro_basicwater, pro_safewater) +
                    te(pro_basicwater, pro_basicsanitation) +
                    te(pro_basicwater, hdi) +
                    te(pro_safewater, pro_basicsanitation) +
                    te(pro_safewater, hdi) +
                    te(pro_basicsanitation, hdi),
                  data = seaa_df, method = "REML")
summary(model_gam_inte)
```
```{r}
# Compare AIC values
AIC(model_lm1, model_lm_inte, model_gam1, model_gam_inte)
```
It can be seen that GAM model with two way interaction performs best.

### Longitudinal analysis with Mixed-effect models

```{r}
# make sure to deal with NA data
seaa_df <- seaa_df %>%
  drop_na(tbi,  pro_basicwater, pro_safewater, pro_basicsanitation, pro_safesanitation, hdi, year, country)

# Model Mixed-Effects: TBI is predicted by predictors with random effect per each country 
model_me <- lmer(tbi ~ pro_basicwater + pro_safewater + pro_basicsanitation + pro_safesanitation + hdi + year + (1 | country),  # Random intercept dressed variation in each country
                   data = seaa_df, REML = FALSE)

# Ringkasan hasil model
summary(model_me)
```

Interpretation: The predictors have no linear relationship with Tuberculosis Incidence. Therefore, we would like to perform nonlinear relationship to decide further action and analysis.

***Check Nonlinear relationship***

```{r}

# Reshape the dataset to long format
asia_africa_long <- seaa_df %>%
  pivot_longer(cols = c( pro_basicwater, pro_safewater, pro_basicsanitation, 
                        pro_safesanitation,hdi),
               names_to = "Predictor", values_to = "Value")

# Plot nonlinear relationships using LOESS smoothing
nonlinear <- ggplot(asia_africa_long, aes(x = Value, y = tbi)) +
  geom_point(alpha = 0.5) +  # Scatter points
  geom_smooth(method = "loess", color = "red") +  # LOESS curve
  facet_wrap(~ Predictor, scales = "free_x") +  # Separate plots for each predictor
  theme_minimal() +
  labs(title = "Nonlinear Relationship between TBI and Predictors",
       x = "Predictor Value",
       y = "TBI")

nonrelationshiplot_location <- here( "results", "figures", "analysis", "nonlinear.png") # to set up location for the pictures created 
ggsave(filename = nonrelationshiplot_location, plot=nonlinear, width = 12, height = 12, units = "in", dpi = 300) # save the pictures created 
plot(nonlinear)
```

It can be seen that some variables show u-shaped (quadratic), however some variables looks more complex. In this case, we are going to try to fit two models, first model will fit Mixed effects model with Polynomial Terms, and addictive mixed model (GAMM).

I will perform two steps here: 1. Data transformation, then looking at possible model, 2. Directly performing Mixed effects model with polynomial terms, and Performing GAMM

```{r}

seaa_df<- seaa_df %>%
  mutate(
    log_tbi = log(tbi + 1),
    sqrt_basicwater = sqrt(pmax(pro_basicwater, 0)), 
    sqrt_safewater = sqrt(pmax(pro_safewater, 0)),
    poly_basicsanitation = as.numeric(poly(pro_basicsanitation, 2, raw = TRUE)[, 1]),
    poly_safesanitation = as.numeric(poly(pro_safesanitation, 2, raw = TRUE)[, 1])
  )

```

Visualize the linearity test after transfomation
```{r}
seaa_transformed <- seaa_df %>%
  dplyr::select(log_tbi, poly_basicsanitation, poly_safesanitation,
                sqrt_basicwater, sqrt_safewater, hdi) %>%
  pivot_longer(cols = -log_tbi, names_to = "Predictor", values_to = "Value")

colnames(seaa_transformed)

# Plot after transformation
plot_transformed <- ggplot(seaa_transformed, aes(x = Value, y = log_tbi)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "loess", color = "red") +  # Use linear model for comparison
  facet_wrap(~ Predictor, scales = "free_x") +
  theme_minimal() +
  labs(title = "Transformed Predictor Relationships with Log(TBI)",
       x = "Transformed Predictor Value", y = "Log(TBI)")

transformedplot_location <- here("results", "figures",  "analysis", "plot_transformed.png") # to set up location for the pictures created 
ggsave(filename = transformedplot_location, plot=plot_transformed, width = 12, height = 12, units = "in", dpi = 300) # save the pictures created 
plot(plot_transformed)
```
*Performing Generalized Additive Models (GAM)*
```{r}

#  Fit the Bayesian GAMM
model_brms_full <- brm(
  formula = tbi ~ 
    s(pro_basicwater) + 
    s(pro_safewater) + 
    s(pro_basicsanitation) + 
    s(pro_safesanitation) + 
    s(hdi) + 
    s(year) + 
    (1 | country),  # random intercept for longitudinal structure

  data = seaa_df,
  family = negbinomial())

```

```{r}
# Model summary
summary(model_brms_full)

# Check model convergence and diagnostics
plot(model_brms_full)           # smooth terms & convergence
pp_check(model_brms_full)       # posterior predictive check
conditional_effects(model_brms_full)  # visualize effects

# Check random effects
ranef(model_brms_full)

# Save fitted values or residuals if needed
seaa_df$fitted_tbi <- fitted(model_brms_full)[, "Estimate"]

```

```{r}
model_brms_full <- brms(
  formula = tbi ~ 
    s(pro_basicwater) + 
    s(pro_safewater) + 
    s(pro_basicsanitation) + 
    s(pro_safesanitation) + 
    s(hdi) + 
    s(year) + 
    (1 | country),

  data = seaa_df,
  family = negbinomial(),

  control = list(adapt_delta = 0.99),  # Fix for divergent transitions

  chains = 4,
  cores = 4,
  iter = 4000
)

)

```

```{r}

wash_vars <- seaa_df[, c("pro_basicwater", "pro_safewater", "pro_basicsanitation", "pro_safesanitation")]
pca_result <- prcomp(wash_vars, scale. = TRUE)

# Add first component to your data
seaa_df$PC1 <- pca_result$x[, 1]

model <- gamm4(
  tbi ~ s(PC1) + s(hdi) + s(year),
  random = ~(1 | country),
  family = nb(),  
  data = seaa_df
)

# Define the model formula with smooth terms
formula <- tbi ~ s(hdi) + s(pro_basicwater) + s(pro_safewater) + s(pro_basicsanitation) + s(pro_safesanitation) + s(year)

# Fit the GAMM with a negative binomial family to account for overdispersion
gamm_model <- gamm(formula, random = list(country = ~1), family = nb(), data = seaa_df)

# Extract the design matrix and response variable
X <- model.matrix(gamm_model$gam)
y <- seaa_df$tbi

# Apply elastic net regularization (alpha = 0.5 for elastic net, 0 for ridge, 1 for lasso)
cv_fit <- cv.glmnet(X, y, alpha = 0.5, family = "poisson")

# Obtain the coefficients at the optimal lambda
coef(cv_fit, s = "lambda.min")

```

after performing data transformation, the data is stil not linear, therefore, Geralized additive mixed morel, which can handle non linearity and account mixed effect of the ciuntries.

```{r}
#convert country into factors

seaa_df$country <- as.factor(seaa_df$country)

model_gamm <- gamm(
  tbi ~ 
    s(hdi) + 
    s(pro_safewater) + 
    s(pro_basicsanitation),
  random = list(country = ~1),
  data = seaa_df,
  method = "REML"
)

model_gamm4 <- gamm4(
  log_tbi ~ 
    s(hdi) +
    s(sqrt_basicwater) +
    s(sqrt_safewater) +
    s(poly_basicsanitation) +
    s(poly_safesanitation),
  random = ~(1 | country),
  data = seaa_df
)

summary(model_gamm4$gam)  # for fixed effects (smooth terms)
summary(model_gamm4$mer)  # for random effects
plot(model_gamm4$gam, pages = 1)

summary(model_gamm_trans$gam)
plot(model_gamm_trans$gam)

```

```{r}
AIC(model1)
AIC(model2)
AIC(model3)
AIC(model4)

BIC(model1)
BIC(model2)
BIC(model3)
BIC(model4)
```

```{r}
# Create a data frame for model comparison
model_comparison <- data.frame(
  Model = c("Model 1: Main Predictors", "Model 2: Including Covariates", "Model 3: Significant variables only", "Model 4: GAMM"),
  AIC = c(AIC(model1), AIC(model2), AIC(model3), AIC(model4)),
  BIC = c(BIC(model1), BIC(model2), BIC(model3), BIC(model4))
)

# Print the table
print(model_comparison)

library(here)

# Define the file location using here()
criteria_location <- here("code", "analysis-code", "tables", "criteria_output.rds")

# Save the model
saveRDS(model_comparison, file = criteria_location)

# Save as a nicely formatted text file
sink("model_comparison.txt")
print(model_comparison)
sink()

```

Model4 (GAM) Performs better.

Lets visualize model4

```{r}

# Define the file path using here()
plot_location <- here("code", "analysis-code", "figures", "gam_diagnostics.jpg")

# Open JPEG device
jpeg(plot_location, width = 2000, height = 1500, res = 300)

# Arrange the plots in a grid (2 rows, 3 columns)
par(mfrow = c(2, 3))  
plot(model4$gam, pages = 1)  # Generate all GAM diagnostic plots

# Close the graphics device to save the file
dev.off()

# Confirm file saved
print(paste("Plot saved at:", plot_location))

```

```{r}
# Create summary table
gam_summary <- summary(model4$gam)

# Extract relevant statistics
gam_table <- data.frame(
  Term = rownames(gam_summary$s.table),
  EDF = gam_summary$s.table[, 1],
  F_Value = gam_summary$s.table[, 3],
  P_Value = gam_summary$s.table[, 4]
)

# Define the file location using here()
summary_location <- here("code", "analysis-code", "tables", "summary_Gamm.rds")

# Save the model
saveRDS(gam_table, file = summary_location)
```

After transforming variables. They look more linear. Now GLM will be performed

***GLMs***

```{r}
glm_model <- glm(log_tbi ~ log_gdp + log_total_greenhouses + log_pm2.5 +
                      poly_cleanfuel + poly_basicsanitation + poly_safesanitation +
                      sqrt_basicwater + sqrt_safewater + hdi,
                 data = asia_africa, family = gaussian)
summary(glm_model)

```

\*\*\*General Addictive Model (GAM)

```{r}
library(mgcv)

gam_model <- gam(log_tbi ~ s(poly_basicsanitation) + s(poly_safesanitation) +
                      s(sqrt_basicwater) + s(sqrt_safewater) + s(hdi),
                 data = seaa_df, family = gaussian)
summary(gam_model)

```

Interpretation:

Generalized Additive Model (GAM) explains 94.5% of the deviance in log-transformed tuberculosis incidence (Log(TBI)), which suggests an excellent fit. Below is a step-by-step breakdown of the interpretation.

Overall Model Performance Adjusted R² (R-sq.(adj) = 0.917) - This means 91.7% of the variability in Log(TBI) is explained by the predictors. - A very strong model since it explains most of the variance.

Deviance Explained = 94.5%

-   Similar to R² but for GAM, indicating excellent model fit.

Generalized Cross-Validation (GCV) = 0.0685

-   Lower GCV indicates better predictive performance.
-   This suggests that the model generalizes well to new data.

![](images/clipboard-4125844072.png)

Now, I want to remove insignificant variables

```{r}
gam_optimized <- gam(log_tbi ~ s(log_gdp) + s(log_total_greenhouses) + s(log_pm2.5) + 
                         s(poly_cleanfuel) + s(sqrt_basicwater) + 
                         s(sqrt_safewater) + s(hdi),
                     data = asia_africa, family = gaussian)
summary(gam_optimized)

```

Comparing GAM models

```{r}
AIC(gam_model, gam_optimized)
BIC(gam_model, gam_optimized)

```

Visualize significant predictors

```{r}
plot(gam_model, pages = 1, shade = TRUE)

```

```{r}
ggplot(data = asia_africa, aes(x = exp(log_tbi), y = exp(predict(gam_model, type = "response")))) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(title = "Actual vs. Predicted TBI",
       x = "Actual TBI", y = "Predicted TBI")

```

***Comparing the model***

```{r}
AIC(glm_model, gam_model)
BIC(glm_model, gam_model)
```

After comapring the models, based on AIC and BIC, GAM model is more appropriate

Check Paryial effects of Predictors

```{r}
plot(gam_model, pages = 1, shade = TRUE)

```

```{r}
predictions <- predict(gam_model, newdata = asia_africa, type = "response")
head(predictions)

```

Compare actual vs predicted values

```{r}
ggplot(data = asia_africa, aes(x = log_tbi, y = predictions)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(title = "Actual vs. Predicted Log(TBI)", x = "Actual Log(TBI)", y = "Predicted Log(TBI)")

asia_africa$predicted_tbi <- exp(predict(gam_model, newdata = asia_africa, type = "response"))

ggplot(asia_africa, aes(x = tbi, y = predicted_tbi)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(title = "Actual vs. Predicted TBI",
       x = "Actual TBI", y = "Predicted TBI")

```

Interpretation:

The points are generally close to the line, indicating that the model is accurate.

```{r}

# Step 1: Ensure all numeric variables are correctly formatted
asia_africa <- asia_africa %>%
  mutate(across(c(tbi, pro_cleanfuel, pro_basicwater, pro_safewater, 
                  mean_exposure_pm2.5, total_greenhouses, pro_basicsanitation, 
                  pro_safesanitation, gdp, hdi), as.numeric)) %>%
  drop_na()  # Remove rows with missing values to avoid ggplot errors

asia_africa <- asia_africa %>%
  mutate(poly_greenhouses = as.numeric(poly(total_greenhouses, degree = 2, raw = TRUE)[,1]))

# Ensure it's a tibble
asia_africa <- as_tibble(asia_africa)

# Select variables (excluding categorical ones)
asia_africa_selected <- asia_africa %>%
  select(-all_of(c("country", "region", "year")))

# Reshape the dataset
asia_africa_long <- asia_africa_selected %>%
  pivot_longer(cols = -tbi, names_to = "predictor", values_to = "value")

# Print first few rows
print(head(asia_africa_long))

# Step 2: Reshape the dataset for visualization
asia_africa <- as_tibble(asia_africa)
asia_africa_long <- asia_africa %>%
  select(-c(country, region, year)) %>%  # Use `-c()` for multiple column exclusion
  pivot_longer(cols = -tbi, names_to = "predictor", values_to = "value")

# Step 3: Check if pivot_longer() worked correctly
print(head(asia_africa_long))

# Step 4: Generate scatterplots with loess smoothing to detect nonlinear patterns
plot_nonlinear <- ggplot(asia_africa_long, aes(x = value, y = tbi)) +
  geom_point(alpha = 0.5) +  # Scatterplot points
  geom_smooth(method = "loess", color = "red") +  # Loess smoothing for nonlinear trends
  facet_wrap(~predictor, scales = "free") +  # Create separate plots for each predictor
  theme_minimal()

# Step 5: Display the plot
print(plot_nonlinear)

nonlinear_location <- here( "code","analysis-code", "figures", "nonlinear.png") # to set up location for the pictures created 
ggsave(filename = nonlinear_location, plot=plot_nonlinear, width = 12, height = 12, units = "in", dpi = 300) # save the pictures created 
plot(combined_plot)
```

```{r}
glm_model <- glm(tbi ~ gdp + hdi + pro_basicsanitation + pro_basicwater + 
                 pro_safesanitation + pro_safewater + pro_cleanfuel + 
                 mean_exposure_pm2.5 + total_greenhouses, 
                 family = poisson(link = "log"), data = asia_africa)
summary(glm_model)

plot(glm_model)

```

Q-Q plot of residuals from the Poisson GLM suggests that the model is not fitting the data well. The deviation from the straight line, especially at higher quantiles (outliers labeled 78, 79, 80), indicates overdispersion—where the variance of tbi is larger than the mean, violating the Poisson assumption.

```{r}
dispersion_test <- sum(residuals(glm_model, type="pearson")^2) / glm_model$df.residual
dispersion_test  # If > 1.5, overdispersion exists

```

since the result is very high, which should be close to 1. We need a model that can hanlde overdispersion

```{r}
library(MASS)
glm_nb <- glm.nb(tbi ~ gdp + hdi + pro_basicsanitation + pro_basicwater + 
                 pro_safesanitation + pro_safewater + pro_cleanfuel + 
                 mean_exposure_pm2.5 + total_greenhouses, data = asia_africa)
summary(glm_nb)

```

Interpretation:

1.  Overdispersion Check The estimated Theta (4.246, SE = 0.430) suggests that overdispersion is present but is well-accounted for by the Negative Binomial model. The Residual Deviance (196.14) is much lower than the Null Deviance (409.01), meaning the model fits the data better than a null model. Dispersion parameter = 1, confirming that the Negative Binomial model is appropriate.

The predictors that significantly impact TBI (tbi) are:

-   hdi (-7.642, p \< 2e-16) Higher HDI is associated with lower tbi cases.
-   pro_basicsanitation (0.00577, p = 0.0334) Increased access to basic sanitation is linked to a slight increase in tbi.
-   pro_basicwater (0.0318, p = 9.42e-16) → Higher access to basic water is positively associated with tbi.
-   pro_safewater (-0.00856, p = 0.0414) → Increased safe water access reduces tbi.
-   pro_cleanfuel (-0.005999, p = 0.0137) → More use of clean fuels is linked to lower tbi.
-   mean_exposure_pm2.5 (-0.02698, p \< 2e-16) → Higher PM2.5 exposure reduces tbi (unexpected; may need further investigation).

Data Transfomation

```{r}
asia_africa <- asia_africa %>%
  mutate(log_tbi = log(tbi + 1),
         log_cleanfuel = log(pro_cleanfuel + 1),
         log_basicwater = log(pro_basicwater + 1),
         log_basicsanitation = log(pro_basicsanitation + 1),
         log_safewater = log(pro_safewater + 1),
         log_safesanitation = log(pro_safesanitation + 1),
         poly_greenhouses = poly(total_greenhouses, 2))  # 2nd-degree polynomial

```

```{r}
library(mgcv)

gam_model <- gam(tbi ~ s(pro_cleanfuel) + s(pro_basicwater) + s(pro_safewater) +
                   s(mean_exposure_pm2.5) + s(total_greenhouses) +
                   s(pro_basicsanitation) + s(pro_safesanitation) +
                   s(gdp) + s(hdi), data = asia_africa, method = "REML")

summary(gam_model)
gamplot<- plot(gam_model, pages = 1)

```

```{r}
# Calculate correlations
cor(asia_africa$tbi, asia_africa$gdp)
cor(asia_africa$tbi, asia_africa$mean_exposure_pm2.5)
cor(asia_africa$tbi, asia_africa$hdi)
cor(asia_africa$tbi, asia_africa$pro_cleanfuel)
cor(asia_africa$tbi, asia_africa$pro_basicwater)
cor(asia_africa$tbi, asia_africa$pro_safewater)
cor(asia_africa$tbi, asia_africa$pro_basicsanitation)
cor(asia_africa$tbi, asia_africa$pro_safesanitation)
cor(asia_africa$tbi, asia_africa$mean_exposure_pm2.5)

```

Spearman's Rank Correlation (Recommended) Spearman correlation is a non-parametric method that measures monotonic relationships (whether the variables move together, not necessarily linearly).

```{r}

# Compute Spearman correlation matrix
cor_matrix_spearman <- cor(asia_africa[, c("tbi","gdp", "hdi", "pro_cleanfuel", "pro_safewater", 
                                           "pro_basicsanitation", "pro_safesanitation", "mean_exposure_pm2.5", "total_greenhouses")], method = "spearman", use = "complete.obs")

# Melt the matrix into long format for ggplot

cor_melted <- melt(cor_matrix_spearman)

# Plot heatmap
plot5 <- ggplot(cor_melted, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
  geom_text(aes(label = round(value, 2)), size = 4) +  # Show correlation values
  theme_minimal() +
  labs(title = "Spearman Correlation Heatmap", x = "", y = "")
matrix_location <- here( "code","eda-code", "figures", "correlation_matrix.png") # to set up location for the pictures created 
ggsave(filename = matrix_location, plot=plot5, width = 12, height = 12, units = "in", dpi = 300) # save the pictures created 
plot(plot5)
```

```{r}

#| label: fig-Spearman's Correlation Matrix 
#| fig-cap: "Spearman's Correlation Matrix "
#| echo: FALSE
knitr::include_graphics(here("code", "eda-code", "figures", "correlation_matrix.png"))
```

I will performed longitudinal data analysis

```{r}
library(car)
#vif(lm(tbi ~ gdp + hdi + pro_cleanfuel + pro_safewater + pro_basicsanitation + 
       pro_safesanitation + mean_exposure_pm2.5 + total_greenhouses, data = pdata)) # The error might be occure due to multicolinearity

#alias(lm(tbi ~ gdp + hdi + pro_cleanfuel + pro_safewater + pro_basicsanitation + 
         pro_safesanitation + mean_exposure_pm2.5 + total_greenhouses, data = asia_africa))

stepwise_model <- stepAIC(lm(tbi ~ gdp + hdi + pro_cleanfuel + pro_safewater + pro_basicsanitation + 
                             pro_safesanitation + mean_exposure_pm2.5 + total_greenhouses, 
                             data = asia_africa), direction = "both") # drop variable with multicolinearity

summary(stepwise_model)

```

***Performing longitudinal analysis***

Test for mixed or fixed effect using Hausman Test

```{r}
pdata <- pdata.frame(asia_africa, index = c("country", "year"))

fe_model <- plm(tbi ~ pro_basicsanitation + mean_exposure_pm2.5, 
                data = pdata, model = "within")
re_model <- plm(tbi ~ pro_basicsanitation + mean_exposure_pm2.5, 
                data = pdata, model = "random")

hausman_test <- phtest(fe_model, re_model)
print(hausman_test)

```

Based on the Hausman Test, fixed effect is not consistent. Therefore, I will use mixed effect test.

```{r}
summary(re_model)
```

Interpretation: ***1. variance components Analyisis***

88% dari total variasi berasal dari perbedaan antar negara, menunjukkan bahwa perbedaan karakteristik tiap negara sangat berpengaruh terhadap insiden TBC.

12% berasal dari variasi idiosinkratik, atau faktor yang berubah dari waktu ke waktu dalam suatu negara.

***2. Theta***

Rata-rata theta = 0.8868, menunjukkan bahwa sebagian besar variasi dalam data disebabkan oleh perbedaan antar negara. Theta yang mendekati 1 berarti bahwa individu (negara) memiliki efek spesifik yang kuat dibandingkan variasi waktu.