### 2025-04-27 project-review-Murtaza Yaqubi.qmd
Project review by Murtaza Yaqubi

---
title: Project Review Template 
author: Murtaza Yaqubi
date: 04/25/2025
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: The effects of Water and Sanitation on Tuberculosis Incidence in South East Asian and African countries

Name of project author(s): Muhammad Nasir

Name of project reviewer: Murtaza Yaqubi

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The context of the project is described very well. The manuscript provides a comprehensive background, explaining that tuberculosis (TB) remains a major global health issue, particularly in South-East Asia and Africa. It outlines the biological and social determinants of TB and highlights that while environmental factors like climate and air pollution have been studied extensively, the role of water and sanitation is less explored. The project is clearly placed within existing research, referencing multiple relevant studies and reports (including WHO data and other recent papers). The motivation for the project is strong and well-articulated: it aims to fill a notable gap in the literature by investigating how water and sanitation access might influence TB incidence, using updated and longitudinal data. It is also clear what new information the project hopes to provide—namely, a better understanding of environmental and development factors in TB burden.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The research question is explicitly stated: to assess whether access to water and sanitation impacts TB incidence. Clear null and alternative hypotheses are provided for each indicator. The link between the research questions and the data is well explained.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data sources (World Bank, WHO, UNICEF) are reliable and clearly mentioned. Variables are listed and described, though a slightly more detailed data dictionary would further assist understanding.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data wrangling is well performed, with careful cleaning, merging, and preparation steps. Exploratory Data Analysis (EDA) includes country-level trends, boxplots of TBI, and multivariable time trends. Some plots, like the correlation heatmap and spaghetti plots of TB incidence, could benefit from deeper interpretation and explicit linking back to the research questions. Tables and datasets can be accessed by making slight modifications to the file path in eda.qmd and edacode.R.

### Summary assessment
*  some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Analysis using Generalized Additive Mixed Models (GAMM) is highly appropriate, capturing non-linear trends and random country effects. Model selection using RMSE and AIC is properly performed. Testing machine learning models like XGBoost to validate results adds robustness. Overall, analysis methods are strong and well-justified.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Results are clearly presented using plots and tables. Figures are generally clean and well-labeled, though occasional improvements to plot titles and clearer faceting could further enhance clarity.

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion is thorough, connecting findings back to hypotheses and literature. Strengths, limitations, and policy implications are addressed thoughtfully. Interpretation of non-significant findings for sanitation is also appropriately cautious.

### Summary assessment
* strong, complete and clear discussion

## Further comments

This project demonstrates excellent understanding of TB determinants and appropriate use of complex modeling techniques.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is very well-structured. Code, results, and data are organized into logical, clearly labeled folders. However, there certain file and folder that are not used and shold be removed.

### Summary assessment
* mostly clear, but some confusing parts (e.g. eda-v2.qmd and supplement folder)

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Documentation is generally very strong, with clear explanations in the manuscript and good code comments. Some exploratory sections could use slightly more detailed annotations for clarity.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The project is fully reproducible. Paths are clearly managed with here(), and minimal manual intervention is required.

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The study is thorough, addressing multiple models, interaction effects, and machine learning validation. Alternative approaches and sensitivity checks (e.g., model tuning) are thoughtfully considered.

### Summary assessment
* strong level of thorougness

## Further comments

This is an outstanding project overall. The attention to model selection, reproducibility, and thorough discussion are commendable. Small additions like a fuller codebook and slightly deeper EDA interpretation would elevate it even more.